Escaping RGBland: Selecting Colors for Statistical Graphics

### 2008-10-22 pkg/paper/Palettes/CSDA.Rnw
small enhancements to CSDA paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@46 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

To overcome these drawbacks, various color spaces have been suggested
that more closely map to the perception dimensions, the most prominent of which are
the CIELUV and CIELAB spaces developed by the \cite{color:CIE:2004}. 
These color spaces are still not perfect---such an ideal is never
truly attainable because color perception will always
differ between individuals---but they offer a significant
improvement over HSV, as Figure~\ref{fig:wheel} illustrates.

@@

Qualitative palettes are sets of colors for depicting different categories,
i.e., for coding a categorical variable. Usually, these should give the same
perceptual weight to each category so that no group is perceived to be larger or more
important than any other one. Typical applications of qualitative palettes in
statistics would be bar plots \citep[see][]{color:Ihaka:2003}, pie charts
or highlighted mosaic displays.

@@

The simplest solution to this task is to employ light/dark contrasts, i.e., rely
on the most basic perceptual axis. The interestingness can be coded
by an increasing amount of gray corresponding to decreasing luminance in HCL space:
  \[ (H, 0, 90 - i \cdot 60), \]
where the hue $H$ used does not matter, chroma is set to 0 (i.e., no color), and 
luminance ranges in $[30, 90]$ avoiding the extreme colors white ($L = 100$) and 
black ($L = 0$). Instead of going linearly from light to dark gray, luminance
could also be increased nonlinearly, e.g., by some function $i' = f(i)$ that
controls whether luminance is increased quickly with intensity or not. We found $f(i) = i^p$ 
to be a convenient transformation where the power $p$ can be varied to achieve
different degrees of non-linearity.

>>>>

To overcome these drawbacks, various color spaces have been suggested
that more closely map to the perception dimensions, the most prominent of which are
the CIELUV and CIELAB spaces developed by the \cite{color:CIE:2004}. 
These color spaces are still not perfect---such an ideal is never
truly attainable because color perception will always
differ between individuals---but they facilitate color matching for humans
\citep{color:Zhang+Montag:2006} and offer a significant
improvement over HSV, as Figure~\ref{fig:wheel} illustrates.

@@

Qualitative palettes are sets of colors for depicting different categories,
i.e., for coding a categorical variable. Usually, these should give the same
perceptual weight to each category so that no group is perceived to be larger or more
important than any other one. Typical applications of qualitative palettes in
statistics would be bar plots \citep[see][]{color:Ihaka:2003}, pie charts
or highlighted mosaic displays (see Section~\ref{sec:illustrations}).

@@

The simplest solution to this task is to employ light/dark contrasts
\citep{color:Ware:1988,color:Levkowitz+Herman:1992}, i.e., rely
on the most basic perceptual axis. The interestingness can be coded
by an increasing amount of gray corresponding to decreasing luminance in HCL space:
  \[ (H, 0, 90 - i \cdot 60), \]
where the hue $H$ used does not matter, chroma is set to 0 (i.e., no color), and 
luminance ranges in $[30, 90]$ avoiding the extreme colors white ($L = 100$) and 
black ($L = 0$). Instead of going linearly from light to dark gray, luminance
could also be increased nonlinearly, e.g., by some function $i' = f(i)$ that
controls whether luminance is increased quickly with intensity or not. We found $f(i) = i^p$ 
to be a convenient transformation where the power $p$ can be varied to achieve
different degrees of non-linearity.
